\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[round]{natbib}
\usepackage{marginnote}
\usepackage{array,warpcol}
\usepackage[acronym,section]{glossaries}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{subfig}


\usepackage[pagebackref=true]{hyperref}

\makeglossaries
\include{Sources/TexFiles/los}
\glossarystyle{tree} 
\renewcommand{\glspostdescription}{.\bigskip}
\bibliographystyle{abbrvnat}
\title{Why, Oh! Why?}
\author{Murukesh Mohanan\\133059001}

\begin{document}
\begin{abstract}
This project aims to implement True Modal Control, whereby every 
controllable mode of a dynamic system will be modelled and controlled 
independently. Dynamic systems can be represented by spring-mass-viscous-damper system  via the Lumped Parameter strategy, but we would have to 
decouple the equations so obtained to obtain the individual modes.
Decoupling classically damped systems is easily done, however not much has 
been done in the case of non-classical damping. We aim to make our method as 
general as possible, so we shall focus on general damping in this project. 
\marginnote{Nearly everything is lifted from my BTP report.}
There have been recent advances in this field by 
\citet{GARVEY2002885,GARVEY2002911} and \citet{Chu200896,Chu2008112}. We 
shall review these advances, and build upon them, principally on the work of 
\citet{Chu200896}.
\end{abstract}
\maketitle
\tableofcontents
\listoffigures
\newpage
\section{Eternal damnation}
Insert random metaphysical stuff here.
\section{Non-eternal damnation}
Insert non-random metaphysical stuuf here.
\subsection{Non-eternal non-damnation}
Insert non-random non-metaphysical stuff here.

\paragraph{Contact Bertrand Russell.}
If not, then please submit your geek card and proceed to Processing.
Something's wrong with atl. Abandon latex plugins, ye who enter vim!
\footnote{Why is :MakeLatex using Python? And more python errors \
	when downloaded from the site.}

\begin{enumerate}
\label{list:class}
\item For $j = 1, \dots, \rho$,
\begin{align}
\left\lbrace \mathrm{r}_j, \mathrm{i}_j\right\rbrace \in \mathbf{C}_a 
&\Leftrightarrow \gls{l}_{\mathrm{r}_j}\gls{l}_{\mathrm{i}_j} > 0\\
\left\lbrace \mathrm{r}_j, \mathrm{i}_j\right\rbrace \in \mathbf{C}_f 
&\Leftrightarrow \gls{l}_{\mathrm{r}_j}\gls{l}_{\mathrm{i}_j} < 0
\end{align}
\item For $j = 1, \dots, (n(\gls{ir}) - \rho)/2$,
\begin{align}
\left\lbrace \mathrm{r}_{\rho+2j-1}, \mathrm{r}_{\rho+2j}\right\rbrace \in \mathbf{C}_b
&\Leftrightarrow \gls{l}_{\mathrm{r}_{\rho+2j-1}}\gls{l}_{\mathrm{i}_{\rho+2j-1}} < 0\\
\left\lbrace \mathrm{r}_{\rho+2j-1}, \mathrm{r}_{\rho+2j}\right\rbrace \in \mathbf{C}_d
&\Leftrightarrow \gls{l}_{\mathrm{r}_{\rho+2j-1}}\gls{l}_{\mathrm{i}_{\rho+2j-1}} > 0
\end{align}
\item For $j = 1, \dots, (n(\gls{ii}) - \rho)/2$,
\begin{align}
\left\lbrace \mathrm{i}_{\rho+2j-1}, \mathrm{i}_{\rho+2j}\right\rbrace \in \mathbf{C}_c
&\Leftrightarrow \gls{l}_{\mathrm{i}_{\rho+2j-1}}\gls{l}_{\mathrm{i}_{\rho+2j-1}} < 0\\
\left\lbrace \mathrm{i}_{\rho+2j-1}, \mathrm{r}_{\rho+2j}\right\rbrace \in \mathbf{C}_e
&\Leftrightarrow \gls{l}_{\mathrm{i}_{\rho+2j-1}}\gls{l}_{\mathrm{i}_{\rho+2j-1}} > 0
\end{align}
\end{enumerate}

\include{Sources/TexFiles/literature}
\section{Dunno.}
\begin{itemize}
	\item Movement of armies:
	\begin{itemize}
		\item Over land:	Slower, delicate units such as siege weapons and 
		monks need to be protected. Exploit the speed of mounted units to 
		ward off attacks on the main army.
		\item Over sea:	Transport ships have limited capacity and units 
		in ships cannot defend themselves. If the enemy is across a body 
		of water, one must set up a base on enemy territory. 
	\end{itemize}
	\item Replenishment of armies after battles and skirmishes:
	\begin{itemize}
		\item Maintaining a level of resources, to allow for 
		quick replenishment.
		\item Ensuring proper timing of arrival of reinforcements.
	\end{itemize}
	\item Retreating under fire while minimizing losses.
\end{itemize}
\section{The table that you requested, milord.}
I hope milord is pleased with my offering.\footnote{Lifted from \href{http://en.wikibooks.org/wiki/LaTeX/Tables}{Wikibooks.}}

\begin{tabular}{ | l | l | l | p{5cm} |}
    \hline
    Day & Min Temp & Max Temp & Summary \\ \hline
    Monday & 11C & 22C & A clear day with lots of sunshine.  
    However, the strong breeze will bring down the temperatures. \\ \hline
    Tuesday & 9C & 19C & Cloudy with rain, across many northern regions. Clear spells
    across most of Scotland and Northern Ireland,
    but rain reaching the far northwest. \\ \hline
    Wednesday & 10C & 21C & Rain will still linger for the morning.
    Conditions will improve by early afternoon and continue
    throughout the evening. \\
    \hline
\end{tabular}

\include{Sources/TexFiles/stable_filtes}

\newpage
\bibliography{Sources/references}
\printglossary[title={List of Symbols},toctitle={List of Symbols}]
\printglossary[type=\acronymtype] 
\end{document}
