\begin{abstract}
This project aims to implement True Modal Control, whereby every 
controllable mode of a dynamic system will be modelled and controlled 
independently. Dynamic systems can be represented by spring-mass-viscous-damper system  via the Lumped Parameter strategy, but we would have to 
decouple the equations so obtained to obtain the individual modes.
Decoupling classically damped systems is easily done, however not much has 
been done in the case of non-classical damping. We aim to make our method as 
general as possible, so we shall focus on general damping in this project. 
There have been recent advances in this field by 
\citet{GARVEY2002885,GARVEY2002911} and \citet{Chu200896,Chu2008112}. We 
shall review these advances, and build upon them, principally on the work of 
\citet{Chu200896}.

We shall attempt decoupling of coupled second-order dynamic equations.
We will take a general quadratic pencil and systematically show how to 
obtain an equivalent pencil involving diagonal matrices. For this we shall 
use the concepts of \glsentryfirstplural{lam} and \glspl{spe}. 
The method described in \citet{Chu200896} will adapted to handle cases of 
singular system matrices, with an example of this case presented. The 
transformations so derived will all be real-valued, keeping in mind that 
most controllers lack support for complex arimthetic. Then we shall discuss 
Automorphic (reflexive) \glspl{spe}, which enable us to develop stable modal 
filters. 

There are infinitely many Automorphic \glspl{spe}, and we shall attempt to 
ascertain the properties of those which result in stable filters. Exhaustive, sequential evaluation of all grid points as well as Monte Carlo searches have been made for stable filters. The study will include both 2-~and 3-~dimensional systems. Plots of the \glspl{cpc} of 
the stable filters provided insight about the space of acceptable   
automorphic \glspl{spe}. Systems in $\mathbb{R}^{2\times 2}$ and 
$\mathbb{R}^{3\times 3}$ have been examined, since higher order systems are 
difficult to visualise. A mixed distribution method has been proposed to increase the efficiency of Monte Carlo searches for stable filters.
\end{abstract}